Methods for Analyzing Models of Rod-Shaped Bacteria

### 2025-07-26 paper/main.tex
split applications part into multiple subsections

% --------------------------------------------------------------------------------------------------
\subsection{Estimating Mechanical Parameters of Growing \textit{E.Coli.} on Agar}
We will now focus on using the presented methods for parameter estimation via microscopic images.
Our initial state is given by TODO~\cite{}.% TODO
We use our extraction algorithm to obtain the initial positions of our cells.
From there, we supply the model with parameters such that we can make a prediction for the final
positions of the bacteria.
We chose a time-interval which does not contain any division events.
By doing so, we can utilize the simpler and more performant position comparison scheme for our cost
function.
Another important detail is the individuality of the agents.
In our case, this means that each agent might have different parameters associated to it.
For this dataset, we assumed that all interaction parameters which specify the interaction
potential are identical across all agents but growth rates and the physical radii of the cells are
individual and we also optimize them.
We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
which were individual for each cell, namely the radii $r_i$ and growth rates $\mu_i$.
We also included the damping constant $\lambda$ and removed any stochastic motion contributions
($D=0$).
This leads to a total of $4+2\times 6=16$ parameters.
To estimate these parameters we used a latin-hypercube sweep across the parameter space and
performed a local minimization with the Nelder-Mead method (TODO citations).% CITATIONS

>>>>

%###################################################################################################
\section{Parameter Estimation of \textit{E.Coli.} growing on Agar}
% --------------------------------------------------------------------------------------------------
\subsection{Data Preparation}
TODO: Decide if this should be Supplement
\begin{enumerate}
    \item Have masks of data; reference image segmantation tools; or do it by hand
    \item ensure that colors are matching for bacteria between time-steps
    \item clean borders of images if needed
    \item know domain size; otherwise use pixels
\end{enumerate}

% --------------------------------------------------------------------------------------------------
\subsection{Non-Individual Treatment}
% --------------------------------------------------------------------------------------------------
\subsection{Individual Treatment}
\label{subsec:parameter-estimation-individual-treatment}
\begin{figure}[H]
    \centering
    \includegraphics[width=0.33\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-individual.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/rod-lengths-average.pdf}%
    \includegraphics[width=0.33\textwidth]
        {figures/crm_estimate_params/IWF-Goettingen/parameter-distribution.pdf}%
    \caption{}
    \label{fig:estimated-growth-rates}
\end{figure}
% --------------------------------------------------------------------------------------------------
\subsection{With Division}
We will now focus on using the presented methods for parameter estimation via microscopic images.
Our initial state is given by TODO~\cite{}.% TODO
We use our extraction algorithm to obtain the initial positions of our cells.
From there, we supply the model with parameters such that we can make a prediction for the final
positions of the bacteria.
We chose a time-interval which does not contain any division events.
By doing so, we can utilize the simpler and more performant position comparison scheme for our cost
function.
Another important detail is the individuality of the agents.
In our case, this means that each agent might have different parameters associated to it.
For this dataset, we assumed that all interaction parameters which specify the interaction
potential are identical across all agents but growth rates and the physical radii of the cells are
individual and we also optimize them.
We chose to optimize for 3 interaction parameters of the Mie potential $n,m,V_0$ and two parameters
which were individual for each cell, namely the radii $r_i$ and growth rates $\mu_i$.
We also included the damping constant $\lambda$ and removed any stochastic motion contributions
($D=0$).
This leads to a total of $4+2\times 6=16$ parameters.
To estimate these parameters we used a latin-hypercube sweep across the parameter space and
performed a local minimization with the Nelder-Mead method (TODO citations).% CITATIONS
